---
title: "Process Leary and Petchey 2009"
author: "Owen Petchey"
date: "6/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)
```


```{r}
rm(list = ls())
library(tidyverse)
library(here)
source(here("../../r/functions.r"))
```

# Import and check datasets

```{r}
study_id <- "leary_and_petchey_2019"
```

```{r}
## check correct folder locations
temp <- stringr::str_split(here(), "/")[[1]]
ifelse(temp[length(temp)] == study_id, "File locations correct.",
   "here() is not returning the correct location.")
```


```{r}
trait <- read_csv(here("data/derived1/trait_data.csv"))
composition <- read_csv(here("data/derived1/community_composition_data.csv"))
community <- read_csv(here("data/derived1/community_data.csv"))
environment <- read_csv(here("data/derived1/environment_data.csv"))
```

# Any trait transformations

```{r}
trait <- trait %>%
  mutate(trait_value = ifelse(trait_name == "carrying capacity",
                           log10(trait_value),
                           trait_value))
```


# Derivative based response diversity

Fit response curves separately for each species and each environmental variable

```{r}
response_models <- trait %>%
  nest_by(species_id, trait_name, environment_name) %>%
  mutate(models = list(mgcv::gam(trait_value ~ s(environment_value, k = 5),
                              data = data))) %>%
  select(-data)
```

Get the derivative for each environmental condition experienced by each species in each community

```{r}
comp_env <- full_join(composition, environment) %>%
  group_by(community_id, species_id, environment_name) %>%
  mutate(new_data = list(data.frame(environment_value = environment_value))) %>%
  select(-environment_value, -environment_time) %>%
  unique() %>%
  full_join(response_models) %>%
  rowwise() %>%
  mutate(derivative = list(gratia::derivatives(models,
                                               newdata = new_data))) %>%
  unnest(derivative) %>%
  select(-new_data, -models) %>%
  rename(environment_value = data)
```

Calculate derivative based response diversity for each community and environmental value, and then take the average across environmental values

```{r}
deriv_resp_div <- comp_env %>%
  group_by(study_id, community_id, trait_name, environment_name, environment_value) %>%
  summarise(temp_mean_nss = resp_div(derivative, sign_sens = FALSE),
            temp_mean_ss = resp_div(derivative, sign_sens = TRUE)) %>%
  group_by(study_id, community_id, trait_name, environment_name) %>%
  summarise(deriv_nss_respdiv = mean(temp_mean_nss),
            deriv_ss_respdiv = mean(temp_mean_ss)) %>%
  pivot_longer(names_to = "community_metric_name",
               values_to = "community_metric_value",
               cols = c("deriv_nss_respdiv", "deriv_ss_respdiv"))
```

# Correlation based response diversity

Get average trait value for each species, trait, environment combination

```{r}

cor_resp_div <- trait %>%
  group_by(study_id, species_id, trait_name, environment_name, environment_value) %>%
  summarise(trait_value = mean(trait_value)) %>%
  right_join(composition) %>%
  right_join(environment) %>%
  pivot_wider(names_from = species_id, names_prefix = "sp_",
              values_from = trait_value) %>%
  group_by(study_id, community_id, trait_name, environment_name) %>%
  do(mean_cor = get_cor(.)) %>%
  mutate(community_metric_name = "cor_respdiv",
         community_metric_value = mean_cor[[1]]) %>%
  select(-mean_cor)
```

# Make final dataset

```{r}
all_respdiv <- bind_rows(cor_resp_div, deriv_resp_div) %>%
  pivot_wider(names_from = community_metric_name,
              values_from = community_metric_value)
all_community <- community %>%
  pivot_wider(names_from = comm_metric_name,
              values_from = comm_metric_value)
all_comm_data <- full_join(all_respdiv, all_community)

write_csv(all_comm_data, here("data/derived2/all_comm_data.csv"))
```


