---
title: "Process Leary and Petchey 2009"
author: "Owen Petchey"
date: "6/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)
```


```{r}
rm(list = ls())
library(tidyverse)
library(here)
source(here("../../r/functions.r"))
```

# Import and check datasets

```{r}
study_name <- "leary_and_petchey_2019"
```

```{r}
## check correct folder locations
temp <- stringr::str_split(here(), "/")[[1]]
ifelse(temp[length(temp)] == study_name, "File locations correct.",
   "here() is not returning the correct location.")
```


```{r}
trait <- read_csv(here("data/derived1/trait_data.csv"))
composition <- read_csv(here("data/derived1/community_composition_data.csv"))
community <- read_csv(here("data/derived1/community_data.csv"))
environment <- read_csv(here("data/derived1/environment_data.csv"))
```

```{r}
studies_dataset_specs <- read_csv(here("../../overall/data/dataset_specifications.csv"))
```

## Trait data

```{r}
var_name_check <- check_variable_names(studies_dataset_specs, trait, "trait")
var_name_check
```

## Composition data

```{r}
var_name_check <- check_variable_names(studies_dataset_specs, composition, "composition")
var_name_check
```

## Community data

```{r}
var_name_check <- check_variable_names(studies_dataset_specs, community, "community")
var_name_check
```

## Environment data

```{r}
var_name_check <- check_variable_names(studies_dataset_specs, environment, "environment")
var_name_check

dataset <- environment
dataset_name_oi <- "environment"

```


# Any trait transformations

```{r}
trait <- trait %>%
  mutate(trait_value = ifelse(trait_name == "carrying capacity",
                           log10(trait_value),
                           trait_value))
```


# Derivative based response diversity

Note that here we calculate the response diversity as the average of the response diversity at the three temperatures experience. In another analysis we did calculate this based on the average response diversity at many levels across the corresponding temperature range. The two methods gave quite strongly correlated results.

Fit response curves separately for each species and each environmental variable

```{r}
response_models <- trait %>%
  nest_by(species_name, trait_name, environment_name) %>%
  mutate(models = list(mgcv::gam(trait_value ~ s(environment_value, k = 5),
                              data = data))) %>%
  select(-data)
```

Get the derivative for each environmental condition experienced by each species in each community

```{r}
comp_env <- full_join(composition, environment) %>%
  group_by(community_name, species_name, environment_name) %>%
  mutate(new_data = list(data.frame(environment_value = environment_value))) %>%
  select(-environment_value, -environment_time) %>%
  unique() %>%
  full_join(response_models) %>%
  rowwise() %>%
  mutate(derivative = list(gratia::derivatives(models,
                                               newdata = new_data))) %>%
  unnest(derivative) %>%
  select(-new_data, -models) %>%
  rename(environment_value = data)
```

Calculate derivative based response diversity for each community and environmental value, and then take the average across environmental values

```{r}
deriv_resp_div <- comp_env %>%
  group_by(study_name, community_name, trait_name, environment_name, environment_value) %>%
  summarise(temp_mean_nss = resp_div(derivative, sign_sens = FALSE),
            temp_mean_ss = resp_div(derivative, sign_sens = TRUE)) %>%
  group_by(study_name, community_name, trait_name, environment_name) %>%
  summarise(deriv_nss_respdiv = mean(temp_mean_nss),
            deriv_ss_respdiv = mean(temp_mean_ss)) %>%
  pivot_longer(names_to = "community_metric_name",
               values_to = "community_metric_value",
               cols = c("deriv_nss_respdiv", "deriv_ss_respdiv"))
```

# Correlation based response diversity

Note that this method is different from that used in the original publication. Here we calculate the average trait at each environment level, and then calculate the correlation of those averages. In the original publication we calculated the average correlation of every possible pairing of the replicates.

Get average trait value for each species, trait, environment combination:

```{r}

cor_resp_div <- trait %>%
  group_by(study_name, species_name, trait_name, environment_name, environment_value) %>%
  summarise(trait_value = mean(trait_value)) %>%
  right_join(composition) %>%
  right_join(environment) %>%
  pivot_wider(names_from = species_name, names_prefix = "sp_",
              values_from = trait_value) %>%
  group_by(study_name, community_name, trait_name, environment_name) %>%
  do(mean_cor = get_cor(.)) %>%
  mutate(community_metric_name = "cor_respdiv",
         community_metric_value = mean_cor[[1]]) %>%
  select(-mean_cor)
```

# Make final dataset

```{r}
all_respdiv <- bind_rows(cor_resp_div, deriv_resp_div) %>%
  pivot_wider(names_from = community_metric_name,
              values_from = community_metric_value)
all_community <- community %>%
  pivot_wider(names_from = community_metric_name,
              values_from = community_metric_value)
all_comm_data <- full_join(all_respdiv, all_community)

comp_4join <- composition %>%
  group_by(community_name) %>%
  summarise(composition = paste(species_name, collapse="-"))
env_4join <- environment %>%
  group_by(community_name, environment_name) %>%
  summarise(environment_values = paste(environment_value, collapse="-"))

all_comm_data <- all_comm_data %>%
  full_join(comp_4join) %>%
  full_join(env_4join)

write_csv(all_comm_data, here("data/derived2/all_comm_data.csv"))
```

# Check some stuff

All evaluations of code chunks set to false.

```{r eval = false}
rm(list = ls())
library(tidyverse)
library(here)
source(here("../../r/functions.r"))
```

```{r eval = FALSE}
new <- read_csv(here("data/derived2/all_comm_data.csv"))
old <- read_csv(here("data/derived2/check_comm_data.csv")) %>%
  mutate(microcosm_ID = str_replace_all(microcosm_ID, "M-", "community-"))
```

```{r eval = FALSE}
new_r <- new %>%
  select(community_name, trait_name,
         cor_respdiv, deriv_nss_respdiv, deriv_ss_respdiv,
         cv_new = cv) %>%
  filter(trait_name == "intrinsic growth rate") %>%
  arrange(community_name) %>%
  select(-trait_name)
new_K <- new %>%
  select(community_name, trait_name,
         cor_respdiv, deriv_nss_respdiv, deriv_ss_respdiv,
         cv_new = cv) %>%
  filter(trait_name == "carrying capacity") %>%
  arrange(community_name) %>%
  select(-trait_name)

old_r <- old %>%
  select(community_name = microcosm_ID,
         r_cor, cv,
         r_dissimilarity, r_divergence)
old_K <- old %>%
  select(community_name = microcosm_ID,
        K_cor = k_cor, cv,
        K_dissimilarity, K_divergence)

compare_r <- full_join(new_r, old_r)
compare_K <- full_join(new_K, old_K)
```

```{r eval = FALSE}
compare_r %>%
  select(-community_name) %>%
GGally::ggpairs()
```

```{r eval = FALSE}
compare_K %>%
  select(-community_name) %>%
GGally::ggpairs()
```

