---
title: "Dataset specifications"
subtitle: "Benchmarking response diversity"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(here)
```


```{r}
dataset_specs <- read_csv(here("overall/data/dataset_specifications.csv"))
```

# Benchmarking datasets

## 1. Studies metadata

The file `data/overall/studies_metadata.csv` contains metadata about each study and a `study_id` that is then used in all datasets for an individual study.

Contains the following variables, though could contain more, e.g. location:

```{r}
dataset_specs %>%
  filter(dataset_name == "studies_metadata") %>%
  select(-dataset_name) %>%
  DT::datatable()

```


## 2. Metrics metadata

The file `overall/data/metrics_metadata.csv` contains metadata about the stability and the response diversity metrics used. Contains the following variables:

```{r}
dataset_specs %>%
  filter(dataset_name == "metrics_metadata") %>%
  select(-dataset_name) %>%
  DT::datatable()
```



# Study datasets



## 1. Trait data

Trait data of some kind is required to calculate the diversity of responses among the set of organisms in a community. The trait information may often be at the species level, such that the trait data is in a species by trait matrix (or in long format, as described below).

The trait information should be in a dataset with the following variables:

```{r}
dataset_specs %>%
  filter(dataset_name == "trait") %>%
  select(-dataset_name) %>%
  DT::datatable()

```




## 2. Community composition data

A list of the species that occur in each community. A required dataset.

```{r}
dataset_specs %>%
  filter(dataset_name == "composition") %>%
  select(-dataset_name) %>%
  DT::datatable()

```

## 3. Community data

Community level data, one value of each metric for each community.


```{r}
dataset_specs %>%
  filter(dataset_name == "community") %>%
  select(-dataset_name) %>%
  DT::datatable()

```
## 4. Environment data

Environmental conditions experienced by each community, an optional dataset.

```{r}
dataset_specs %>%
  filter(dataset_name == "environment") %>%
  select(-dataset_name) %>%
  DT::datatable()

```

## Community stability and response diversity data

This dataset is produced by processing of the previous three or four.

```{r}
dataset_specs %>%
  filter(dataset_name == "all_community_data") %>%
  select(-dataset_name) %>%
  DT::datatable()

```